---
title: "Practical Mashine Learning "
author: "Olga Smirnov"
date: "October 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(randomForest)
library(caret)
library(e1071)

```

## Overview

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

The goal of this project is to build machine learning algorithm to predict activity quality ( classe) from activity monitor


## Data
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

The training data are available here https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

and the test data here 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

#### We download the two datasets

```{r }

#read in training and testing data
train <- read.csv("training.csv", na.strings=c("NA","#DIV/0!",""))
test <- read.csv("testing.csv", na.strings=c("NA","#DIV/0!",""))
dim(train)
```
The training data has 19622 observations and 160 features, and the distribution of the five measured classe A,B,C,D,E is:
```{r}

summary(train$classe)                      #this is the outcome we want to predict

##names(train)
str(train)
```
#### Clining Data

First 7 coloms are not predictors
```{r}
train <-train[,8:length(train)]
```

remove variables with near zero variance
```{r} 
NZV <-nearZeroVar(train)
train <- train[, -NZV]
```
Remove variables with more than 75% or NAs
```{r}
testNA <-sapply(train,function(x)mean(is.na(x))) >0.75
train <-train[,testNA==F]
```
#### Partitioning the training set

We separate our training data into a training set and a validation set
```{r}


inTrain <- createDataPartition(y=train$classe, p=0.6, list=FALSE)
myTrain <- train[inTrain, ]
myValid <- train[-inTrain, ]
dim(myTrain)

dim(myValid)
```
## Random Forest Model

Build the model
```{r}
set.seed(12345)

RFModel <- randomForest(classe~., data = myTrain)
print(RFModel)

RFPredict <- predict(RFModel, myValid )
cm<-confusionMatrix(myValid$classe, RFPredict)
print(cm)
cm$overall['Accuracy']
```

The cross validation accuracy is 99.4% and the out-of-sample error is therefore 0.6%  so our model performs rather good.

## Run model against Testing data
```{r}
predicttest <- predict(RFModel, test)
print(predicttest)
```
